### 2016-11-25 thesis/intro/intro.tex
tmp commit

Metastasis is a multi-step process that requires the selection and survival of cancer cell(s) that is able to metastasise (``seed''), and is also compatible with the specific target tissue where they prosper (``soil'') \citep{Talmadge2010}.
The downregulation or loss of function of E-cadherin, a protein important in cell-to-cell connection and adherence, is known to trigger tumour invasion and metastasis through the reversible process of \gls{emt} \citep{Hanahan2011,Kalluri2009}.
Once mobile and invasive, these metastatic cancer cell is able to circulate through the vascular system and eventually invade the target organ, where it must undergo \gls{met} and colonise the site, and eventually metastasise from this site \citep{Hanahan2011,Kalluri2009}.

@@

There are two pathways of inflammation that contribute to the cancer survival and progression \citep{Mantovani2008}.
The intrinsic pathway is caused by oncogenic characteristics of the cancer cells, where these cells produce inflammatory signals themselves and promote advantageous characteristics for the cancer cells, such as increased proliferation and survival \citep{Mantovani2008}.
On the other hand, the external pathway is caused by infectious or inflammatory conditions in the tumour microenvironment that surround the tumour and assist in the growth of the tumour \citep{Mantovani2008}.
There are two key transcription factors that connect the two pathways of inflammation that ultimately leads to the prommotion  of many, if not all, of the hallmarks discussed so far: \gls{nfkb} and \gls{stat3} \citep{Mantovani2008}.

@@

>>>>

Metastasis is a multi-step process that requires the selection and survival of cancer cell(s) that is able to metastasise (``seed''), and is also compatible with the specific target tissue where they prosper (``soil'') \citep{Talmadge2010}.
The downregulation or loss of function of E-cadherin, a protein important in cell-to-cell connection and adherence, is known to trigger tumour invasion and metastasis through the reversible process of \gls{emt} \citep{Hanahan2011,Kalluri2009}.
Once mobile and invasive, these metastatic cancer cell is able to circulate through the vascular system and eventually invade the target organ, where it must undergo \gls{met} and colonise the site \citep{Hanahan2011,Kalluri2009}.

@@

There are two pathways of inflammation that contribute to the cancer survival and progression \citep{Mantovani2008}.
The intrinsic pathway is caused by oncogenic characteristics of the cancer cells, where these cells produce inflammatory signals themselves and promote advantageous characteristics for the cancer cells, such as increased proliferation and survival \citep{Mantovani2008}.
On the other hand, the external pathway is caused by infectious or inflammatory conditions in the tumour microenvironment that surround the tumour and assist in the growth of the tumour \citep{Mantovani2008}.
There are two key transcription factors that connect these two pathways of inflammation that ultimately leads to the prommotion  of many, if not all, of the hallmarks discussed so far: \gls{nfkb} and \gls{stat3} \citep{Mantovani2008}.

@@

\noindent
In order to grow and proliferate, cancer cells must adjust their energy metabolism to match their energy requirement \citet{Hanahan2011}.
There are two ways in which a cell can produce energy: anaerobic respiration, where the cell uses glycolysis to rapidly generate \gls{atp} and lactate; and aerobic respiration where, in the presence of oxygen, the cell utilises the electron transport chain and \gls{atp} synthase in the mitochondria to efficiently produce \gls{atp}.
However, even with abundant oxygen in the cell, \citet{Wardburg1956} found that cancer cells utilised the glycolytic pathway instead of the more efficient aerobic respiration, which was later termed ``aerobic glycolysis'' \citep{Hanahan2011}.

It seems counterintuitive to limit its full potential of creating \gls{atp} by utilising glycolysis over electron transport chain, as the latter produces almost 18 times more \gls{atp} \citep{Hanahan2011, VanderHeiden2009}.
However, it is hypothesised  that this apprent inefficiency helps the cell to generate and direct the molecules required for biosynthetic pathways during cell growth and division \citep{Cairns2011,VanderHeiden2009}.
By deliberately relying purely on glycolysis (known as the ``Wardberg effect'') and slowing the conversion of \gls{pep} to pyruvate, the cell is able to shunt the carbon molecules into reaction pathways that produce constituents used to synthesise macromolecules, such as \acrshort{dna} and lipids \citep{Cairns2011,VanderHeiden2009}.

The fact that the deregulation of energy metabolism is mediated somewhat by the genes involved in many of the other hallmarks (such as p53 and Myc) suggests that the disruption in energy metabolism may be a phenotype caused in conjunction with those core hallmarks \citep{Hanahan2011}.
